---
title: "MCF_Weather_Processing"
author: "Chih-Yu"
date: "2025-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(data.table)
library(tidyverse)
```

## 1. Daily average temperature 1981-2019

```{r}
main_dir <- "C:/ECCC/MCF/MCF-improvement/Inputs/SLC_Weather_WRH"
station_folders <- list.dirs(main_dir, recursive = FALSE)

station_results <- list()

for (station_path in station_folders) {
  station_name <- basename(station_path)
  message("Processing: ", station_name)

  file_list <- list.files(station_path, pattern = "\\.txt$", full.names = TRUE)

  file_matrices <- lapply(file_list, function(f) {
    dt <- tryCatch(fread(f, header = FALSE), error = function(e) return(NULL))
    if (is.null(dt) || nrow(dt) < 365 || ncol(dt) < 3) return(NULL)
    as.matrix(dt[1:365, .(V2, V3)])
  })

  file_matrices <- Filter(Negate(is.null), file_matrices)
  if (length(file_matrices) == 0) next

  array_data <- simplify2array(file_matrices)
  avg_matrix <- apply(array_data, c(1, 2), mean, na.rm = TRUE)
  final_vector <- rowMeans(avg_matrix)

  station_results[[station_name]] <- final_vector
}

# Combine into one data frame
station_df <- bind_rows(station_results, .id = "Station")

# Add month labels (ignoring leap years)
days_in_months <- c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
month_labels <- rep(1:12, times = days_in_months)
station_df$Month <- month_labels
station_df <- station_df %>% relocate(Month)

saveRDS(station_df, "../Inputs/Daily Weather.RDS")
```

## 2. Define Utility Functions for Seasonal Logic
This section includes logic for identifying manure removal months based on temperature thresholds.(Hung et al., 2022)

A. Early spring removal was set to themonth,which had the first occurrence
of average daily air temperature above 5 째C. Exceptions were granted for
locations in British Columbia, where the removal was set to April, otherwise
the warmer climate could result in removals in January (five locations)
or March (35 locations).
B. Late spring removal was one month later than the early spring removal.
C. Early fall removal was set to the month 45 d before the first day in fall to
reach 0 째C. In the case of three removals per year, the summer removal
was in August and the months of early spring and late fall were determined
as mentioned above.
D. Late fall removal was set to the first fall month with an average daily air
temperature of 0 째C. If the average air temperature remained above 0 째C
for the entire fall period, removal was set to November.



```{r }
months_vector <- station_df$Month

get_first_month_over_5 <- function(values, months, min_month = 4) {
  idx <- which(values >= 5)[1]
  if (is.na(idx)) return(min_month)
  return(max(months[idx], min_month))
}

get_late_spring_month <- function(values, early_spring_day, months) {
  idx <- early_spring_day + 30
  if (idx > length(months)) idx <- length(months)
  return(months[idx])
}

get_early_fall_month <- function(values, months) {
  idx_start <- which(months >= 9)[1]
  idx_cold <- which(values[idx_start:length(values)] <= 0)[1]
  if (is.na(idx_cold)) return(9)
  idx_cold_absolute <- idx_start + idx_cold - 1
  idx_early_fall <- max(1, idx_cold_absolute - 45)
  return(months[idx_early_fall])
}

get_late_fall_month <- function(values, months) {
  idx_start <- which(months >= 9)[1]
  idx_cold <- which(values[idx_start:length(values)] <= 0)[1]
  if (is.na(idx_cold)) return(11)
  return(months[idx_start + idx_cold - 1])
}

```

## 3. Determine Removal Months for Each Station
We now apply the rules to generate the removal schedule for early/late spring and early/late fall.


```{r }
Month_removal <- tibble(
  Station = colnames(station_df)[-1],
  Early_Spring = NA_integer_,
  Late_Spring = NA_integer_,
  Early_Fall = NA_integer_,
  Late_Fall = NA_integer_
)

for (i in seq_along(Month_removal$Station)) {
  station <- Month_removal$Station[i]
  values <- station_df[[station]]

  idx_early_spring <- which(values >= 5)[1]
  if (is.na(idx_early_spring)) idx_early_spring <- which(months_vector == 4)[1]
  early_spring_month <- max(months_vector[idx_early_spring], 4)

  late_spring_month <- get_late_spring_month(values, idx_early_spring, months_vector)
  early_fall_month <- get_early_fall_month(values, months_vector)
  late_fall_month <- get_late_fall_month(values, months_vector)

  Month_removal$Early_Spring[i] <- early_spring_month
  Month_removal$Late_Spring[i]  <- late_spring_month
  Month_removal$Early_Fall[i]   <- early_fall_month
  Month_removal$Late_Fall[i]    <- late_fall_month
}



#Save results
write_csv(Month_removal, "../Outputs/Month_removal_schedule.csv")
```


